---
title: "Joshua Tree Deposition"
author: "Kym Howo, Alex Irvin, Margaux Sleckman, Caitlin Swalec"
date: "5/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load Packages and Functions

```{r packages}

### Need to install the gt package from github for the results table###
library(devtools)
remotes::install_github("rstudio/gt")
###

library(tidyverse)
library(deSolve)
library(sensitivity)
library(gt)

```

```{r functions}
source("WetDeposition.R")
source("DryDeposition.R")
source("HumanPopulationGrowth.R")
source("TransportEmissions.R")
source("PollutionTransport.R")
```

#### Population Growth ODE Submodel

The first part of the model is finding how the LA population will grow over the next 100 years. This will ultimately determine how many vehicles are present in LA in the emissions phase of the model. In LA, the initial population is 10,105,518 and the growth rate is 0.13%. The carrying capacity is set to 100,000,000, which ends up not mattering given the slow growth rate of the population (final pop is 11.3 million in 100 years). This is the ODE submodel for this project.

```{r ode}

#Set up the time frame
time_frame = data.frame(time=seq(from=1,to=100))

#Set initial parameters of population size, growth rate, and carrying capacity
parameters = list(p0=10105518,r=0.0013,K=100000000)

#Run ODE model to find human population over the time frame
result = ode(parameters$p0, time_frame$time, HumanPopGrowth, parameters)

#Create dataframe to pass into next submodel
LA_population = data.frame(result)
colnames(LA_population) = c("time","pop")

```

#### Emissions Submodel

The second part of the model is the emissions phase, where automobile emissions dispense NOx into the atmosphere. The output of this model is the automobile nitrogen emissions in LA for each year in the sample, which will be used as input into the pollution transport model. 


```{r emissions}

#Run emissions submodel and create dataframe with emissions of Nitrogen (in grams) per year
LA_auto_emissions = transport_emissions_LA(LA_population)
colnames(LA_auto_emissions) = c("time","pop","nitrogen")

```

#### Transport Submodel

This submodel evaluates how the nitrogen emissions in LA will translate to nitrogen in the air in Joshua Tree. This is done through a linear regression of historical data for air quality in Los Angeles and Coachella Valley (closest South Coast Air Quality Management District monitoring location to Joshua Tree). 

```{r message = FALSE}

#Read in historic NOx data from LA and Joshua Tree. City 1 is LA, City 2 is Joshua Tree.
historic_data = read_csv("SCAQMD data raw JT.csv")
#To demonstrate ability to run with another city's data, try reading in "SCAQMD data raw SBer.csv" where City 1 is LA and City 2 is San Bernardino Valley.

#Run transport submodel and create a dataframe with Joshua Tree nitrogen (in grams) for each time step
JT_nitrogen = air_transport(historic_data, LA_auto_emissions)
JT_nitrogen_df = data.frame(JT_nitrogen[1])
JT_nitrogen[7]
JT_nitrogen[8]

```

The boxplots show that there are no outliers in this data set and the density plots show that the distribution of NOx for both LA and Joshua Tree are relatively normal. The Joshua Tree data starts to show signs of a bimodal distribution, but we decided that it is reasonable to assume normality. The linear regression correlated NOx levels in LA to NOx levels in Joshua Tree. Correlation between these data sets is strong with a correlation factor of 0.94, showing that it is reasonable to assumer linear correlation. The equation for the linear regression produced using the historic data for Joshua Tree and Los Angeles is as follows: NOx level in Joshua Tree = -8.0148 + (NOx level in Los Angeles)*0.6693


#### Deposition Submodel

In this phase, wet and dry deposition is calculated from the Nitrogen that is left in the atmosphere after the transport phase. The output of each model is a deposition rate in kg/m2. These rates are then added together and converted to kg/hectare for comparison to the deposition threshold described by Bell and Allen.

```{r}

#Run wet and dry deposition models, which calculate rate of deposition
wet_deposition = WetDeposition(JT_nitrogen_df)
dry_deposition = DryDeposition(JT_nitrogen_df)

#Add wet and dry deposition rates together
total_deposition = (wet_deposition$wet_dep + dry_deposition$dry_dep) * 0.0001

#Create dataframe with the additional deposition rate for each year
deposition_rate_per_year = data.frame(time = wet_deposition$time, additional_dep = total_deposition)

```

The ouput of this model is a dataframe of deposition rates per year. These rates represent the additional nitrogen deposition that results from automobile emissions in LA. 

#### Final Result

After all the models have been run, the results are displayed in a GT table and a ggplot.

```{r result}

#GT Table of added deposition each year
colnames(deposition_rate_per_year) = c("Year","AddedDeposition")
deposition_rate_per_year %>% 
  gt() %>% 
  tab_header(
    title = "Added Deposition in Joshua Tree", # Add a title
    subtitle = "kg/hectare"# And a subtitle
  ) %>%
  fmt_passthrough( # Not sure about this but it works...
    columns = vars() # First column: supp (character)
  )

ggplot(deposition_rate_per_year) + geom_line(aes(x = Year, y = AddedDeposition)) + theme_classic() + ylab("Additional Deposition\n") + xlab("\nYear")

```

As expected, nitrogen deposition increases over time due to increasing automobile emissions.

#### Sensitivity Analysis 

```{r}

#Create two samples for sobel sensitivity with annual precipitation and number of storms changing
nsample=1000
Ps1 = cbind.data.frame(
  p = runif(min=0, max=1, n=nsample))

Ps2 = cbind.data.frame(
  p = runif(min=0, max=1, n=nsample))

Ns1 = cbind.data.frame(
  ns = runif(min=0, max=365, n=nsample))

Ns2 = cbind.data.frame(
  ns = runif(min=0, max=365, n=nsample))

allp1 = cbind.data.frame(Ps1, Ns1)
allp2 = cbind.data.frame(Ps2, Ns2)

#Run sobol sensitivity model
sens_micro=sobol2007(model = NULL, allp1, allp2, nboot = 100)
nsim=nrow(sens_micro$X)

#Set up data frame
res = rep(0,times=nsim)

#Loop through every simulation for wet deposition in year 100
for (i in 1:nsim) {
  tmp = WetDeposition(nitrogen_concentration = JT_nitrogen_df,
                      annual_precip=as.numeric(sens_micro$X[i,c("p")]),
                      num_storms = as.numeric(sens_micro$X[i,c("ns")])) 
  
  res[i] = tmp$wet_dep[100]
} 

#Give our results to sensitivity structure
sens_micro = sensitivity::tell(sens_micro, res)

tmp_sens = cbind.data.frame(sens_micro$X, nitrogen100=sens_micro$y)

par(mfrow=c(1, 2))
ggplot(tmp_sens, aes(p, nitrogen100))+
  geom_point()+
  ggtitle("Sensitivity analysis of wet nitrogen deposition \n based on annual precipitation") +
  labs(x="Annual Precipitation",y="Additional Wet Nitrogen Deposition")
ggplot(tmp_sens, aes(ns, nitrogen100))+
  geom_point()+
  ggtitle("Sensitivity analysis of wet nitrogen deposition \n based on annual precipitation") +
  labs(x="Number of Storms",y="Additional Wet Nitrogen Deposition")

```

This sensitivity analysis shows that annual precipitation has an affect on wet deposition in year 100, but additional wet deposition approaches an upper limit as annual precipitation approaches infinity. At lower levels of precipitation, the level of wet deposition tends to be closer to the upper limit, but at higher levels of precipitation, wet deposition is widely distributed. As the number of storms increase, the level of wet deposition also tends to increase, again with an upper limit on wet deposition. Wet deposition is clustered closely to the upper limit on deposition as the number of storms increases, suggesting that storm intensity (level of precipitation per storm) has a significant impact on wet deposition.


**Citations**
Air Quality data - South Coast Air Quality Management District, Historical Data By Year
http://www.aqmd.gov/home/air-quality/air-quality-data-studies/historical-data-by-year

Deposition threshold for Joshua Tree - Bell, Michael, and Allen, Edith. "Wind, earth, and fire: The impacts of anthropogenic air pollution on soils in Joshua Tree National Park"
https://www.nps.gov/articles/parkscience32_2_64-66_bell_allen_3838.htm
